---
title: "Correlation and Regression"
subtitle: "StatPREP worksheet example"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
knitr::opts_chunk$set(echo = FALSE)

fast_food <- read_csv("https://raw.githubusercontent.com/joeroith/Data/master/Nutritional_Data_for_Fast_Food_2017.csv")
```

## Fast Food Nutrition

### Background and Data

Most of the calories in food come from fats, carbohydrates, and proteins.

The dataset was collected in January of 2017 by looking through online nutritional information provided by fast food restaurant chains. Nutrition data on various burgers, a breaded chicken sandwich, a grilled chicken sandwich, chicken nuggets, french fries, and a chocolate milkshake were collected for each restaurant (when applicable). For each chain the smallest hamburger, the smallest cheeseburger, and a variety of their most well known larger burgers were selected.

```{r}
fast_food

# variable names
names(fast_food)
```

## Plot the Data

Make a scatterplot of the number of calories in a fast food item versus how many grams of carbohydrates it contains.

```{r plot}
ggplot(data = __________, mapping = aes(x = _______, y = _______)) +
  geom_point() +
  labs(title = "...")
```

```{r plot-hint}
ggplot(data = fast_food, mapping = aes(x = carbs, y = calories)) +
  geom_point() +
  labs(title = "A scatterplot of Carbs (g) vs. Calories")
```

#### Question 1

Does there appear to be a relationship between the two variables?

## Summarize the Data

Find the summary statistics for `calories` and `carbs`. Be sure to find means, standard deviations, and the correlation coefficient.

```{r summary}
fast_food %>%
  summarize(...)
```

```{r summary-hint}
fast_food %>%
  summarize(cal_mean = mean(calories),
            cal_sd = sd(calories),
            carb_mean = mean(carbs),
            carb_sd = sd(carbs),
            correlation = cor(calories, carbs)
  )

# You could also use df_stats()
```

#### Question 2

Make a statement about the strength and direction of the relationship between Calories and Carbohydrates.

## Least-squares Regression

### Simple regression

Find the intercept and the slope of the least-squares regression line for `calories` vs. `carbs`.

```{r olsr, exercise = TRUE}
model <- lm(_____ ~ ______, data = ______)
```

```{r olsr-hint}
model <- lm(calories ~ carbs, data = fast_food)
summary(model)
```

#### Question 3

What is the correct interpretation of the intercept of the least-squares regression line?

#### Question 4

What is the correct interpretation of the slope of the least-squares regression line?

## BONUS: Explore through Regression

### Sodium vs. Sugar

Create a scatterplot of `sodium` and `sugar`.

```{r plot2, exercise = TRUE}
ggplot(data = _______, mapping = aes(x = _______, y = ________)) +
  geom_?
```

```{r plot2-hint}
ggplot(data = fast_food, mapping = aes(x = sodium, y = sugars)) +
  geom_point() +
  labs(title = "Scatterplot of Sodium (mg) vs. Sugars (g)")
```

There seems to be a cluster of points throwing off our pattern. Create a new scatterplot and color the points based on `type` of fast-food item. (Add regression lines and facet the plots if you wish as well.)

```{r plot3, exercise = TRUE}

```

```{r plot3-hint}
ggplot(data = fast_food, mapping = aes(x = sodium, y = sugars, color = type)) +
  facet_wrap(~ type) +
  geom_point() +
  geom_smooth(method = "lm")
```
