---
title: "Using the mosaic package"
subtitle: "StatPREP Fort Myers Workshop 2022"
author: "Kathryn Kozak"
date: "5/20/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mosaic)
library(NHANES)
library(mosaicData)
data_NHANES<-NHANES
no_NA_diabetes<-data_NHANES[!is.na(data_NHANES$Diabetes),]
no_NA_education<-data_NHANES[!is.na(data_NHANES$Education),]
```

## Reading in datasets

Reading in a dataset:

```{r}
births_2015<-Births2015 #reads a built in dataset into the global environment
```

```{r}
Affairs<-read.csv("/cloud/project/data/Affairs.csv") #this dataset is uploaded into your files
```

```{r}
Wages<- read.csv("https://krkozak.github.io/MAT160/wages.csv") #Using github to store dataset
```

It helps to know what variables the dataset contains. This command will display the names of the variables in the dataset.

```{r}
names(data_NHANES)
```

IF you want to look at the dataset

```{r}
View(data_NHANES)
```

Sometimes you want to just look at the structure of the dataset, so it is helpful to look at the first 6 rows of a dataset using this command

```{r}
head(data_NHANES)
```

To see the code book on a dataset from a package:

```{r}
?NHANES
```

##mosaic Package

R has packages that can be installed with datasets such as mosaicData, NHANES, gapminder, fivethirtyeight, and many others.

The package mosaic also is a package that makes some graphs and data commands easy to compute. This session will explore some commands in the mosaic package using the NHANES dataset.

##Categorical Data

Make a bar chart of a categorical variable.

```{r}
gf_bar(~Race1, data=data_NHANES)
```

Sometime is it useful to have the bar graph be horizontal.

```{r}
gf_barh(~Race1, data=data_NHANES)
```

We can add a title, color, subtitle, and x and y labels

```{r}
gf_bar(~Race1, data=data_NHANES, title="Race of people who live in the US", fill="purple", xlab="Race", ylab="Number of People")
```

If you would rather have proportions instead of count:

```{r}
gf_props(~Race1, data=data_NHANES, title="Race of people who live in the US", fill="purple", xlab="Race", ylab="Proportion of People")
```

Suppose you want to facet by a categorical variable and display as a stacked bar graph:

```{r}
gf_bar(~Race1, data=no_NA_education, title="Race of people who live in the US", subtitle = "Education Level", fill=~Education, xlab="Race", ylab="Number of People")
```

Display as side by side bars:

```{r}
gf_bar(~Race1, data=no_NA_education, title="Race of people who live in the US", subtitle = "Education Level", fill=~Education, position=position_dodge(), xlab="Race", ylab="Number of People")
```

Creating a Pareto Chart, to go from highest to lowest

```{r}
gf_props(~reorder(Race1,Race1,function(Race1)-length(Race1)), data=data_NHANES, title="Race of people who live in the US", fill="purple", xlab="Race", ylab="Proportion of People")
```

## Quantitative Data

To make a histogram of a qualitative variable.

```{r, warning=FALSE}
gf_histogram(~TotChol, data=data_NHANES, title = "Total Cholesterol of people in US", fill="blue", color="red")
```

To make a histogram of a qualitative variable faceted by a categorical variable.

```{r, warning=FALSE}
gf_histogram(~TotChol|Diabetes, data=no_NA_diabetes, title = "Total Cholesterol of people in US", subtitle = "Person has Diabetes", fill="blue", color="red")
```

To make a density graph of a quantitative variable.

```{r, warning=FALSE}
gf_density(~TotChol|Diabetes, data=no_NA_diabetes,title = "Total Cholesterol of people in US", fill="blue")
```

```{r, warning=FALSE}
gf_density(~TotChol, data=no_NA_diabetes, title = "Total Cholesterol of people in US", fill=~Diabetes)
```

```{r, warning=FALSE}
gf_density(~BPSysAve, data=data_NHANES, fill="purple", title="Blood Pressure", subtitle = "Purple=systolic, and Blue=diastolic")%>%
  gf_density(~BPDiaAve, data=data_NHANES, fill="blue")
```

## Scatterplot and Time Series graphs

Scatterplot of two quantitative variables

```{r}
gf_point(length~width, data=KidsFeet, title="Length versus Width of Kids longer foot")
```

```{r}
gf_point(length~width, data=KidsFeet, color=~domhand, title="Length versus Width of Kids longer foot")
```

Make a scatter plot for number of births each day of the year in 2015 in US.

```{r}
gf_point(births~date, data=Births2015, title="Number of Births in 2015")
```

```{r}
gf_point(births~date, data=Births2015, title="Number of Births in 2015", subtitle="Day of the Week", color=~wday)
```

Create a time series graph for the time in seconds of the 100 m swimming world record faceted by sex.

```{r}
gf_line(time~year, data=SwimRecords, title="temperature", color=~sex)
```

## Statistics

Finding the mean, median, variance, standard deviation, range, and five-number summary of a dataset

```{r}
help(SaratogaHouses)
```

```{r}
df_stats(~price, data=SaratogaHouses, mean)
```

Faceted by a categorical variable

```{r}
df_stats(~price|bedrooms, data=SaratogaHouses, mean)
```

boxplots

```{r}
gf_boxplot(~price|bedrooms, data=SaratogaHouses, title="Prices of Houses Separated by Number of Bedrooms")
```

##Inference

Hypothesis tests and confidence intervals are computed using the same commands.

Proportion test

```{r}
prop.test(outcome~smoker, data=Whickham) #example of survival status whether smoker
```

If want a one-sided test:

```{r}
prop.test(outcome~smoker, data=Whickham, alternative="less") #example of survival status whether smoker
```

A one sample test of the mean:

```{r}
t.test(~alcohol, mu=9.87, data=Alcohol) #comparing the number of liters drunk around the world to US average in 2018 of 9.87 Liters per day.
```

two sample independent test of the difference in means:

```{r}
t.test(age~smoker, data=Whickham) #comparing the age of a person and whether the person smokes or does
```

For a paired sample test, you need to create a difference variable using the mutate command.

linear regression:

```{r}
lm(length~width, data=KidsFeet) #creates a linear model for the length of a kids feet versus the weidth
```

```{r}
gf_point(length~width, data=KidsFeet, title="Length versus Width of Kids longer foot")%>%
gf_lm()
```

```{r}
lm.out<-lm(length~width, data=KidsFeet) #saves the linear model to the variable lm.out
summary(lm.out) #displays summary information about the linear model lm.out
```
